Behavior Driven Development in Haskell

### 2012-05-28 paper.tex
using "evidence".

We can categorize examples/properties into two types: for users and
and for developers.
Examples/properties for users are those which are useful as user documentation.
By contrast, examples/properties for developers are
those which are not useful as user documentation.
A good example of examples/properties for developers
is a corner example registered in a bug tracking system.

BDD tools should be treat both of them and provide functionality
to extract examples/properties for users as user documentation.
If we can archive this, BDD tools can be used for
triple purpose: design, documentation and automatic testing.

@@

To describe examples/properties for developers, we can use Hspec.

@@

\begin{itemize}
\item
    The Haskell community does not have common syntax to describe
    examples/properties with Haddock markup syntax.
\item
    The Haskell community does not have a BDD tool for
    examples/properties for users.
\item
    Hspec does not provide combinators to express expectations with
    BDD terminology
\end{itemize}

@@

To be able to specify examples and properties for users, we propose to add new
markup to Haddock and to use Doctest integrated with QuickCheck. To
describe examples/properties for developers, we propose to use Hspec also
integrated with QuickCheck.

@@

Examples/properties for users should be documented with Haddock markup syntax.
For this purpose, we propose the followings:

@@

\noindent Examples and properties in comments are automatically
tested by the Doctest command. We also provided a library
to do this job. Its implementation will discussed in Section~\ref{sec:doctest}.

@@

\begin{quote}
    \emph{TBD: Behavior should be replaced with examples/properties in this subsection?}
\end{quote}

@@

\noindent The {\tt it} function combines a text description for a behavior and
a property or an example that gives some confidence in that behavior into a
spec.
Examples can be arbitrary HUnit assertions.  Properties must be of type {\tt
Property}; QuickCheck's \texttt{property} function can be used to turn
arbitrary properties inot a \texttt{Property}.

@@

A spec can be run with the \texttt{hspec} function.  When a spec is run, all
properties and examples are evaluated, and failing properties and examples are
reported to the user (FIXME: user does not fit our definition of \emph{user}
here!).

@@

As we described, Haddock version 2.7.2 or earlier provides markup syntax for
code blocks, but does not provide any markup syntax for examples or properties.
We implemented markup support for examples ({\tt >>>}) and it is available in
Haddock version 2.8 or later.
We also extended Haddock to support markup for properties ({\tt prop>}) which
will be available with a future release of Haddock.
Both examples and properties in Haddock comments are pretty-printed when
generating documentation, (e.g. as HTML).

@@

TBD(@sol): how to obtain examples and properties from Haddock markup.
Old version uses Haddock API but now uses GHC API directly. Why?

@@

\noindent Note that the {\tt placeholders} library provide better keyword such as {\tt notImplemented} and {\tt todo}.
Next we write examples/properties for users in Haddock markup:

@@

\noindent We should execute Doctest to ensure no syntax errors.
Also, we specifies examples/properties for developers with Hspec:

>>>>

We can categorize evidences into two types: for users and
and for developers.
evidences for users are those which are useful as user documentation.
By contrast, evidences for developers are
those which are not useful as user documentation.
A good example of evidences for developers
is a corner example registered in a bug tracking system.

BDD tools should be treat both of them and provide functionality
to extract evidences for users as user documentation.
If we can archive this, BDD tools can be used for
triple purpose: design, documentation and automatic testing.

@@

To describe evidences for developers, we can use Hspec.

@@

\begin{itemize}
\item
    The Haskell community does not have common syntax to describe
    evidences with Haddock markup syntax.
\item
    The Haskell community does not have a BDD tool for
    evidences for users.
\item
    Hspec does not provide combinators to express expectations with
    BDD terminology
\end{itemize}

@@

To be able to specify evidences for users, we propose to add new
markup to Haddock and to use Doctest integrated with QuickCheck. To
describe evidences for developers, we propose to use Hspec also
integrated with QuickCheck.

@@

evidences for users should be documented with Haddock markup syntax.
For this purpose, we propose the followings:

@@

\noindent Evidences in comments are automatically
tested by the Doctest command. We also provided a library
to do this job. Its implementation will discussed in Section~\ref{sec:doctest}.

@@

\begin{quote}
    \emph{TBD: Behavior should be replaced with evidences in this subsection?}
\end{quote}

@@

\noindent The {\tt it} function combines a text description for a behavior and
an evidence that gives some confidence in that behavior into a
spec.
Examples can be arbitrary HUnit assertions.  Properties must be of type {\tt
Property}; QuickCheck's \texttt{property} function can be used to turn
arbitrary properties inot a \texttt{Property}.

@@

A spec can be run with the \texttt{hspec} function.  When a spec is run, all
evidences are evaluated, and failing evidences are
reported to the user (FIXME: user does not fit our definition of \emph{user}
here!).

@@

As we described, Haddock version 2.7.2 or earlier provides markup syntax for
code blocks, but does not provide any markup syntax for evidences.
We implemented markup support for examples ({\tt >>>}) and it is available in
Haddock version 2.8 or later.
We also extended Haddock to support markup for properties ({\tt prop>}) which
will be available with a future release of Haddock.
Both examples and properties in Haddock comments are pretty-printed when
generating documentation, (e.g. as HTML).

@@

TBD(@sol): how to obtain evidences from Haddock markup.
Old version uses Haddock API but now uses GHC API directly. Why?

@@

\noindent Note that the {\tt placeholders} library provide better keyword such as {\tt notImplemented} and {\tt todo}.
Next we write evidences for users in Haddock markup:

@@

\noindent We should execute Doctest to ensure no syntax errors.
Also, we specifies evidences for developers with Hspec:
